Flowchart of Current Algorithm

### 2009-04-29 docs/flowchart/flowchart.tex
notes on how to time-center  gamma1bar in advect_base_pres_spherical.  not implemented in code yet

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
\rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:

\begin{color}{red}
%Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.  The pressure update will use $\gammabar^{\inp,n+\myhalf}$.  To compute $\psi^{\outp,n}$, we use $\gammabar^{\inp,n}$.  To compute $\psi^{\outp,n+\myhalf}$, we use $\gammabar^{\inp,n+\myhalf}$.  In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.  For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,\star})/2$.  I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
\end{color}

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{equation}
We time-center the left hand side:
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\end{equation}
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Now we begin the corrector step.
\begin{color}{red} We need to compute 
$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{(2)},p_0^\star,X_k^{(2)})]$ 
and $\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, and then use 
$\gammabar^\dagger$ in the $f^\dagger$ equation. \end{color} 
Using $p_0^\star$, we can compute a more 
accurate approximation to equation~(\ref{eq:frhs}),
\begin{equation}
f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $p_0^\dagger$ on edges.

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}

>>>>

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{equation}
We time-center the left hand side:
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\end{equation}
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Next, we compute a provisional updated $\gammabar$ using
$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$.\\

Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
we can compute a more accurate approximation to equation~(\ref{eq:frhs}),
\begin{equation}
f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, 
$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, 
and the difference is again computed by first putting $p_0^\dagger$ on edges.

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
\rho^{(2)},X_k^{(2)}]$\\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}
